---
title: "320 Final Project"
author: "Belle Schmidt and Rachael Olson"
date: "2025-11-12"
output: html_document
---

```{r}
library(readr)
library(dplyr)
library(pracma)
nfl_team_stats_2002_2024 <- read_csv("C:\\Users\\isabe\\OneDrive\\Big Data Bowl\\TeamRanking\\Schmidt\\nfl_team_stats_2002-2024.csv")
View(nfl_team_stats_2002_2024)

# nfl_team_stats_2002_2024
```

```{r}
nfl_team_stats_2002_2024 <- nfl_team_stats_2002_2024 %>% 
  mutate(away_win_margin = score_away - score_home,
         home_win_margin = score_home - score_away) |> 
  select(season, week, home, away, home_win_margin, score_away, score_home) |> 
  mutate(home = ifelse(home == "Bills", "T1", ifelse(home == "Dolphins", "T2", ifelse(home == "Patriots", "T3", ifelse(home == "Jets", "T4", ifelse(home == "Ravens", "T5", ifelse(home == "Bengals", "T6", ifelse(home == "Browns", "T7", ifelse(home == "Steelers", "T8", ifelse(home == "Texans", "T9", ifelse(home == "Texans", "T9", ifelse(home == "Colts", "T10", ifelse(home == "Jaguars", "T11", ifelse(home == "Titans", "T12", ifelse(home == "Broncos", "T13", ifelse(home == "Chiefs", "T14", ifelse(home == "Raiders", "T15", ifelse(home == "Chargers", "T16", ifelse(home == "Cowboys", "T17", ifelse(home == "Giants", "T18", ifelse(home == "Eagles", "T19", ifelse(home == "Commanders", "T20", ifelse(home == "Bears", "T21", ifelse(home == "Lions", "T22", ifelse(home == "Packers", "T23", ifelse(home == "Vikings", "T24", ifelse(home == "Falcons", "T25", ifelse(home == "Panthers", "T26", ifelse(home == "Saints", "T27", ifelse(home == "Buccaneers", "T28", ifelse(home == "Cardinals", "T29", ifelse(home == "Rams", "T30", ifelse(home == "49ers", "T31", "T32"))))))))))))))))))))))))))))))))) |> 
  mutate(away = ifelse(away == "Bills", "T1", ifelse(away == "Dolphins", "T2", ifelse(away == "Patriots", "T3", ifelse(away == "Jets", "T4", ifelse(away == "Ravens", "T5", ifelse(away == "Bengals", "T6", ifelse(away == "Browns", "T7", ifelse(away == "Steelers", "T8", ifelse(away == "Texans", "T9", ifelse(away == "Texans", "T9", ifelse(away == "Colts", "T10", ifelse(away == "Jaguars", "T11", ifelse(away == "Titans", "T12", ifelse(away == "Broncos", "T13", ifelse(away == "Chiefs", "T14", ifelse(away == "Raiders", "T15", ifelse(away == "Chargers", "T16", ifelse(away == "Cowboys", "T17", ifelse(away == "Giants", "T18", ifelse(away == "Eagles", "T19", ifelse(away == "Commanders", "T20", ifelse(away == "Bears", "T21", ifelse(away == "Lions", "T22", ifelse(away == "Packers", "T23", ifelse(away == "Vikings", "T24", ifelse(away == "Falcons", "T25", ifelse(away == "Panthers", "T26", ifelse(away == "Saints", "T27", ifelse(away == "Buccaneers", "T28", ifelse(away == "Cardinals", "T29", ifelse(away == "Rams", "T30", ifelse(away == "49ers", "T31", "T32"))))))))))))))))))))))))))))))))) |> 
  filter(week %in% 1:18)


Season24 <- nfl_team_stats_2002_2024 %>% 
  filter(season == 2024)

Season23 <- nfl_team_stats_2002_2024 |> 
  filter(season == 2023)

Season22 <- nfl_team_stats_2002_2024 |> 
  filter(season == 2022)

Season21 <- nfl_team_stats_2002_2024 |> 
  filter(season == 2021)

# Thru_Week_6 <- Season24 %>% 
#   filter(week < 6) %>% 
#   dplyr::select(-season, -week, -date, -time_et)
# 
# Thru_Week_6
# 
# Away_Stats <- Thru_Week_6 %>% 
#   group_by(away) %>% 
#   summarise(points_allowed_away = sum(score_home),
#             yards_allowed_away = sum(yards_home),
#             sacks_away = sum(sacks_num_away),
#             turnovers_away = sum(fumbles_away) + sum(interceptions_away),
#             num_away_games = n())
# 
# Home_Stats <- Thru_Week_6 %>% 
#   group_by(home) %>% 
#   summarise(points_allowed_home = sum(score_away),
#             yards_allowed_home = sum(yards_away),
#             sacks_home = sum(sacks_num_home),
#             turnovers_home = sum(fumbles_home) + sum(interceptions_home),
#             num_home_games = n())
# 
# 
# ThruWeek6Totals <- full_join(Away_Stats, Home_Stats, join_by(away == home))

# Season24 %>% 
#   dplyr::select(-season, -date, -time_et, -first_downs_away, -first_downs_home, -first_downs_from_passing_away, -first_downs_from_passing_home, -first_downs_from_rushing_away, -first_downs_from_rushing_home, -first_downs_from_penalty_away, -first_downs_from_penalty_home, -third_down_comp_away, -third_down_comp_home, -third_down_att_away, -third_down_att_home, -fourth_down_att_away, -fourth_down_att_home, -fourth_down_comp_away, -fourth_down_comp_home, -plays_away, -plays_home, -drives_away, -drives_home, -pass_comp_away, -pass_comp_home)

#View(Season24)
```

# Pre Information

AFC East: 
T1 = Bills
T2 = Dolphins
T3 = Patriots
T4 = Jets

AFC North: 
T5 = Ravens
T6 = Bengals
T7 = Browns
T8 = Steelers

AFC South: 
T9 = Texans
T10 = Colts 
T11 = Jaguars
T12 = Titans

AFC West: 
T13 = Broncos
T14 = Chiefs 
T15 = Raiders
T16 = Chargers

NFC East: 
T17 = Cowboys
T18 = Giants
T19 = Eagles
T20 = Commanders

NFC North: 
T21 = Bears
T22 = Lions
T23 = Packers
T24 = Vikings

NFC South: 
T25 = Falcons
T26 = Panthers
T27 = Saints
T28 = Buccaneers

NFC West: 
T29 = Cardinals
T30 = Rams
T31 = 49ers
T32 = Seahawks

# Functions 

# Comparing Rankings 

```{r}

# Exact Comparison
Comparison <- function(ranking1, ranking2){
  ranking1 <- ranking1 |> 
  select(Team) |> 
  rename(Ranking_1 = Team)

  ranking2 <- ranking2 |> 
  select(Team) |> 
  rename(Ranking_2 = Team)

  Ranking_comparison <- cbind(ranking1, ranking2)|> 
    mutate(differs = Ranking_1 != Ranking_2)
  
  return(Ranking_comparison)
}

# Comparison within Three Places
move_3_comparison <- function(ranking1, ranking2) {
  ranking1 <- ranking1 |> #team names and rank nums
    dplyr::select(Team) |> 
    dplyr::mutate(Ranking_1 = dplyr::row_number())

  ranking2 <- ranking2 |> 
    dplyr::select(Team) |> 
    dplyr::mutate(Ranking_2 = dplyr::row_number())
  
  #join rankings by Team
  Ranking_comparison <- dplyr::left_join(ranking1, ranking2, by = "Team")

  #team has or has not moved more than 3 spots
  Ranking_comparison <- Ranking_comparison |>
    dplyr::mutate(moved_3_or_less = abs(Ranking_1 - Ranking_2) <= 3)

  return(Ranking_comparison)
}
```

# Massey's Method General: Least Squares

```{r}
# Function That Makes Initial T Matrix
modify_matrix <- function(input_matrix, data) {
  for (i in 1:nrow(data)){
    col <- data$home[i]
    col_index <- as.numeric(sub("T","", data$home[i]))
    input_matrix[i, col_index] <- 1
    col2 <- data$away[i]
    col_index2 <- as.numeric(sub("T","", data$away[i]))
    input_matrix[i, col_index2] <- -1
    
  
  }
  
  return(input_matrix)
}

# Used For Fixing The Linear Dependence Issue
modify_matrix2 <- function(input_matrix, row) {
  for (i in 1:ncol(input_matrix)){
    input_matrix[row, i] <- 1
    
}
  
  return(input_matrix)
    
}
```

# Keener's Method 

```{r}
A_matrix <- function(input_data) {
  zeros <- matrix(0, ncol = 32, nrow = 32)
  for (i in 1:(nrow(input_data))){
    if ((input_data$home[i] %in% 1:4 && input_data$away[i] %in% 1:4) |
        (input_data$home[i] %in% 5:8 && input_data$away[i] %in% 5:8) |
        (input_data$home[i] %in% 9:12 && input_data$away[i] %in% 9:12) |
        (input_data$home[i] %in% 13:16 && input_data$away[i] %in% 13:16) |
        (input_data$home[i] %in% 17:20 && input_data$away[i] %in% 17:20) |
        (input_data$home[i] %in% 21:24 && input_data$away[i] %in% 21:24) |
        (input_data$home[i] %in% 25:28 && input_data$away[i] %in% 25:28) |
        (input_data$home[i] %in% 29:32 && input_data$away[i] %in% 29:32)){
       for (j in 1:nrow(input_data)){
         if (input_data$home[i] == input_data$away[j] && input_data$away[i] == input_data$home[j]){
           row_index <- input_data$home[i]
          # print(message("The two teams are in the same division. The home team in the first game is", row_index))
           col_index <- input_data$away[i]
           # print(message("The away team is", col_index))
           zeros[row_index, col_index] <- (input_data$score_home[i] + input_data$score_away[j] + 1) / (input_data$score_home[i] + input_data$score_away[i] + input_data$score_home[j] + input_data$score_away[j] + 2)
           # print(message("The aij is", zeros[row_index, col_index]))
           
           zeros[col_index, row_index] <- (input_data$score_away[i] + input_data$score_home[j] + 1) / (input_data$score_home[i] + input_data$score_away[i] + input_data$score_home[j] + input_data$score_away[j] + 2)
         #  print(message("The aji is", zeros[col_index, row_index]))
         }
       }
    }
    else{
    row_index <- input_data$home[i]
    col_index <- input_data$away[i]
    # print(message("These teams are not in the same division. The home team was", row_index))
    # print(message("The away team was", col_index))
    zeros[row_index,col_index] <- (input_data$score_home[i] + 1) / (input_data$score_home[i] + input_data$score_away[i] + 2)
    # print(message("The aij is", zeros[row_index,col_index]))
       zeros[col_index,row_index] <- (input_data$score_away[i] + 1) / (input_data$score_home[i] + input_data$score_away[i] + 2)
   #  print(message("The aji is", zeros[col_index,row_index]))
  }
  }
  return(zeros)
    
}

Skewing_Matrix <- function(input_matrix){
  for (i in 1:nrow(input_matrix)) {
    for (j in 1:ncol(input_matrix)) {
      if (input_matrix[i,j] == 0) {
        input_matrix[i,j] = 1/2
      }
      else{
      input_matrix[i, j] = (1/2) + (((sign(input_matrix[i,j] - (1/2)) * sqrt(abs(2* input_matrix[i,j] - 1))))/(2))
      }
    } 
  }
  return(input_matrix)
}
```

# Season 2024 

```{r}
Week_9_Least_Squares <-Season24 |> 
  mutate(week = as.numeric(week)) |> 
  filter(week < 10)
```

```{r}
NFL_End_Of_Season_Ranking_2024 <- data.frame(Team = c("Lions", "Chiefs", "Vikings", "Eagles","Bills", "Ravens", "Commanders", "Packers", "Chargers", "Broncos", "Texans", "Rams", "Steelers", "Seahawks", "Buccaneers", "Bengals", "Cardinals", "Falcons", "Colts", "Dolphins", "Cowboys", "49ers", "Panthers", "Bears", "Saints", "Jets", "Jaguars", "Raiders", "Patriots", "Browns", "Giants", "Titans"))

NFL_Playoff_Placing <- data.frame(Team = c("Eagles", "Chiefs", "Bills", "Commanders","Lions","Ravens","Rams", "Texans", "Buccaneers","Vikings","Chargers","Steelers","Packers","Broncos"))
```

>> Keener's method had the Eagles as number 2, while the regular NFL rankings had them at number 4 and least squares at number 5; Keeners had chiefs at number 7 though when NFL had then at number 2. Least squares had them at number 10. Bills were ranked 6th in keener, 5th by the NFL, and 4th from least squares. Commanders were number 10 in keeners, 11 for least squares, and 7th in NFL. 

# Week 18 Keener's Method Rankings 2024

```{r}
Season24 |> 
  filter(home %in% c("T1","T2","T3","T4","T5", "T27","T28","T29","T30","T31","T32") & away %in% c("T27","T28","T29","T30","T31","T32", "T1","T2","T3","T4","T5"))

Week18_For_Keener_24 <- Season24 |> 
  mutate(home = as.numeric(sub("T","", home)),
         away = as.numeric(sub("T","", away)))

Initial_Matrix_Week18_24 <- A_matrix(Week18_For_Keener_24) 

Matrix_Week18_24 <- Skewing_Matrix(Initial_Matrix_Week18_24)

eigen_Week18_24 <- eigen(Matrix_Week18_24)

team_ranking_week18_24 <- data.frame(Team = c("Bills", "Dolphins", "Patriots", "Jets", "Ravens", "Bengals", "Browns", "Steelers", "Texans", "Colts", "Jaguars", "Titans", "Broncos", "Chiefs", "Raiders", "Chargers", "Cowboys", "Giants", "Eagles", "Commanders", "Bears", "Lions", "Packers", "Vikings", "Falcons", "Panthers", "Saints", "Buccaneers", "Cardinals", "Rams", "49ers", "Seahawks"),
                         Ranking_Value = eigen_Week18_24$vectors[,1]) |> 
  arrange(Ranking_Value)

team_ranking_week18_24 
```

## Least Squares Ranking Week 18 2024

```{r}
Solutions_week18_24 <- as.matrix(Season24[, "home_win_margin"])

zeros_week18_24 <- matrix(0, ncol = 32, nrow = 272)

T_week18_24 <- modify_matrix(zeros_week18_24, Season24)

NewT_week18_24 <- t(T_week18_24)%*%T_week18_24

NewSolution_week18_24 <- t(T_week18_24)%*%Solutions_week18_24

LinearIndependence_week18_24 <- modify_matrix2(NewT_week18_24, 32)

NewSolution_week18_24[32, 1] <- 0

NFL_Rankings_Week18_24 <- solve(LinearIndependence_week18_24) %*% NewSolution_week18_24

Ranking_Least_SquaresWeek18_24 <- data.frame(Team = c("Bills", "Dolphins", "Patriots", "Jets", "Ravens", "Bengals", "Browns", "Steelers", "Texans", "Colts", "Jaguars", "Titans", "Broncos", "Chiefs", "Raiders", "Chargers", "Cowboys", "Giants", "Eagles", "Commanders", "Bears", "Lions", "Packers", "Vikings", "Falcons", "Panthers", "Saints", "Buccaneers", "Cardinals", "Rams", "49ers", "Seahawks"),
                         Ranking_Value = NFL_Rankings_Week18_24[,1]) |> 
  arrange(-Ranking_Value)

Ranking_Least_SquaresWeek18_24
```

# 2023 NFL Rankings 

>> Chiefs won. 49ers 2nd. Ravens and Lions lost in conference championships. All three rankings had ravens winning the super bowl. Both least squares and keeners had 49ers as runner ups but NFL had the 49ers in 4th. Keeners had the chiefs 5, NFL had them as 7th, and least sqaures had them as 6th. Lions were ranked 6th by keeners, 3rd by the NFL and 7th by least squares. 

```{r}
NFL_End_Of_Season_Ranking_2023 <- data.frame(Team = c("Ravens", "Cowboys", "Lions", "49ers", "Bills", "Browns", "Chiefs", "Dolphins", "Eagles", "Texans", "Rams", "Steelers", "Bengals", "Packers", "Colts", "Jaguars", "Saints", "Seahawks", "Buccaneers", "Broncos", "Raiders", "Falcons", "Bears", "Vikings", "Jets", "Giants", "Titans", "Chargers", "Cardinals", "Patriots", "Commanders", "Panthers"))
```

# Week 18 Keener's Method Rankings 2023

```{r}
Week18_For_Keener_23 <- Season23 |> 
  mutate(home = as.numeric(sub("T","", home)),
         away = as.numeric(sub("T","", away)))

Initial_Matrix_Week18_23 <- A_matrix(Week18_For_Keener_23) 

Matrix_Week18_23 <- Skewing_Matrix(Initial_Matrix_Week18_23)

eigen_Week18_23 <- eigen(Matrix_Week18_23)

team_ranking_week18_23 <- data.frame(Team = c("Bills", "Dolphins", "Patriots", "Jets", "Ravens", "Bengals", "Browns", "Steelers", "Texans", "Colts", "Jaguars", "Titans", "Broncos", "Chiefs", "Raiders", "Chargers", "Cowboys", "Giants", "Eagles", "Commanders", "Bears", "Lions", "Packers", "Vikings", "Falcons", "Panthers", "Saints", "Buccaneers", "Cardinals", "Rams", "49ers", "Seahawks"),
                         Ranking_Value = eigen_Week18_23$vectors[,1]) |> 
  arrange(Ranking_Value)

team_ranking_week18_23 
```

## Least Squares Ranking Week 18 2023

```{r}
Solutions_week18_23 <- as.matrix(Season23[, "home_win_margin"])

zeros_week18_23 <- matrix(0, ncol = 32, nrow = 272)

T_week18_23 <- modify_matrix(zeros_week18_23, Season23)

NewT_week18_23 <- t(T_week18_23)%*%T_week18_23

NewSolution_week18_23 <- t(T_week18_23)%*%Solutions_week18_23

LinearIndependence_week18_23 <- modify_matrix2(NewT_week18_23, 32)

NewSolution_week18_23[32, 1] <- 0

NFL_Rankings_Week18_23 <- solve(LinearIndependence_week18_23) %*% NewSolution_week18_23

Ranking_Least_SquaresWeek18_23 <- data.frame(Team = c("Bills", "Dolphins", "Patriots", "Jets", "Ravens", "Bengals", "Browns", "Steelers", "Texans", "Colts", "Jaguars", "Titans", "Broncos", "Chiefs", "Raiders", "Chargers", "Cowboys", "Giants", "Eagles", "Commanders", "Bears", "Lions", "Packers", "Vikings", "Falcons", "Panthers", "Saints", "Buccaneers", "Cardinals", "Rams", "49ers", "Seahawks"),
                         Ranking_Value = NFL_Rankings_Week18_23[,1]) |> 
  arrange(-Ranking_Value)

Ranking_Least_SquaresWeek18_23
```

# NFL 2022 Rankings

>> Chiefs won the super bowl. Eagles finished 2nd. Bengals and 49ers lost in conference championships. The NFL had the Chiefs winning the super bowl when the keeners ranked them 4 and least squares had them ranked 6th. The eagles were ranked number 2 by the NFL and keeners while least squares had them at 4. The bengals were ranked 6th by the NFL, 5th by the keeners method and least squares. The 49ers were ranked 3rd by keeners, 5th by the NFL, and 2nd by least squares. 

```{r}
NFL_End_Of_Season_Ranking_2022 <- data.frame(Team = c("Chiefs", "Eagles", "Bills", "Vikings", "49ers", "Bengals", "Cowboys", "Ravens", "Chargers", "Giants", "Lions", "Jaguars", "Dolphins", "Steelers", "Seahawks", "Commanders","Packers", "Patriots", "Buccaneers", "Falcons", "Panthers", "Browns", "Saints", "Jets", "Titans", "Raiders", "Broncos", "Rams", "Colts", "Cardinals", "Texans", "Bears"))
```

# Week 18 Keener's Method Rankings 2022

```{r}
Week18_For_Keener_22 <- Season22 |> 
  mutate(home = as.numeric(sub("T","", home)),
         away = as.numeric(sub("T","", away)))

Initial_Matrix_Week18_22 <- A_matrix(Week18_For_Keener_22) 

Matrix_Week18_22 <- Skewing_Matrix(Initial_Matrix_Week18_22)

eigen_Week18_22 <- eigen(Matrix_Week18_22)

team_ranking_week18_22 <- data.frame(Team = c("Bills", "Dolphins", "Patriots", "Jets", "Ravens", "Bengals", "Browns", "Steelers", "Texans", "Colts", "Jaguars", "Titans", "Broncos", "Chiefs", "Raiders", "Chargers", "Cowboys", "Giants", "Eagles", "Commanders", "Bears", "Lions", "Packers", "Vikings", "Falcons", "Panthers", "Saints", "Buccaneers", "Cardinals", "Rams", "49ers", "Seahawks"),
                         Ranking_Value = eigen_Week18_22$vectors[,1]) |> 
  arrange(-Ranking_Value)

team_ranking_week18_22 
```

## Least Squares Ranking Week 18 2022

```{r}
Solutions_week18_22 <- as.matrix(Season22[, "home_win_margin"])

zeros_week18_22 <- matrix(0, ncol = 32, nrow = 271)

T_week18_22 <- modify_matrix(zeros_week18_22, Season22)

NewT_week18_22 <- t(T_week18_22)%*%T_week18_22

NewSolution_week18_22 <- t(T_week18_22)%*%Solutions_week18_22

LinearIndependence_week18_22 <- modify_matrix2(NewT_week18_22, 32)

NewSolution_week18_22[32, 1] <- 0

NFL_Rankings_Week18_22 <- solve(LinearIndependence_week18_22) %*% NewSolution_week18_22

Ranking_Least_SquaresWeek18_22 <- data.frame(Team = c("Bills", "Dolphins", "Patriots", "Jets", "Ravens", "Bengals", "Browns", "Steelers", "Texans", "Colts", "Jaguars", "Titans", "Broncos", "Chiefs", "Raiders", "Chargers", "Cowboys", "Giants", "Eagles", "Commanders", "Bears", "Lions", "Packers", "Vikings", "Falcons", "Panthers", "Saints", "Buccaneers", "Cardinals", "Rams", "49ers", "Seahawks"),
                         Ranking_Value = NFL_Rankings_Week18_22[,1]) |> 
  arrange(-Ranking_Value)

Ranking_Least_SquaresWeek18_22
```

# NFL Rankings 2021

```{r}
NFL_End_Of_Season_Ranking_2021 <- data.frame(Team = c("Packers", "Buccaneers", "Cowboys", "Chiefs", "Rams", "Titans", "Cardinals", "Bills", "Bengals", "Raiders", "Patriots", "49ers", "Steelers", "Colts", "Chargers", "Dolphins","Saints", "Eagles", "Ravens", "Browns", "Vikings", "Falcons", "Broncos", "Seahawks", "Commanders", "Bears", "Panthers", "Texans", "Giants", "Jets", "Lions", "Jaguars"))
```


>> Rams won the super bowl. Bengals finished 2nd. Chiefs and 49ers lost in conference championships. The NFL and keeners method had the rams as 5th while least squares had them 6th. Bengals were ranked 9th by the NFL, they were ranked 11th by keeners, and 12th by least squares. The chiefs were ranked number 4 by the NFL and keeners. Least squares had them at number 5.     

# Week 18 Keener's Method Rankings 2021

```{r}
Week18_For_Keener_21 <- Season21 |> 
  mutate(home = as.numeric(sub("T","", home)),
         away = as.numeric(sub("T","", away)))

Initial_Matrix_Week18_21 <- A_matrix(Week18_For_Keener_21) 

Matrix_Week18_21 <- Skewing_Matrix(Initial_Matrix_Week18_21)

eigen_Week18_21 <- eigen(Matrix_Week18_21)

team_ranking_week18_21 <- data.frame(Team = c("Bills", "Dolphins", "Patriots", "Jets", "Ravens", "Bengals", "Browns", "Steelers", "Texans", "Colts", "Jaguars", "Titans", "Broncos", "Chiefs", "Raiders", "Chargers", "Cowboys", "Giants", "Eagles", "Commanders", "Bears", "Lions", "Packers", "Vikings", "Falcons", "Panthers", "Saints", "Buccaneers", "Cardinals", "Rams", "49ers", "Seahawks"),
                         Ranking_Value = eigen_Week18_21$vectors[,1]) |> 
  arrange(Ranking_Value)

team_ranking_week18_21 
```

## Least Squares Ranking Week 18 2021

```{r}
Solutions_week18_21 <- as.matrix(Season21[, "home_win_margin"])

zeros_week18_21 <- matrix(0, ncol = 32, nrow = 272)

T_week18_21 <- modify_matrix(zeros_week18_21, Season21)

NewT_week18_21 <- t(T_week18_21)%*%T_week18_21

NewSolution_week18_21 <- t(T_week18_21)%*%Solutions_week18_21

LinearIndependence_week18_21 <- modify_matrix2(NewT_week18_21, 32)

NewSolution_week18_21[32, 1] <- 0

NFL_Rankings_Week18_21 <- solve(LinearIndependence_week18_21) %*% NewSolution_week18_21

Ranking_Least_SquaresWeek18_21 <- data.frame(Team = c("Bills", "Dolphins", "Patriots", "Jets", "Ravens", "Bengals", "Browns", "Steelers", "Texans", "Colts", "Jaguars", "Titans", "Broncos", "Chiefs", "Raiders", "Chargers", "Cowboys", "Giants", "Eagles", "Commanders", "Bears", "Lions", "Packers", "Vikings", "Falcons", "Panthers", "Saints", "Buccaneers", "Cardinals", "Rams", "49ers", "Seahawks"),
                         Ranking_Value = NFL_Rankings_Week18_21[,1]) |> 
  arrange(-Ranking_Value)

Ranking_Least_SquaresWeek18_21
```

# Week 9 Keener's Method Rankings

```{r}
Week9_For_Keener <- Week_9_Least_Squares |> 
  mutate(home = as.numeric(sub("T","", home)),
         away = as.numeric(sub("T","", away)))

Initial_Matrix_Week9 <- A_matrix(Week9_For_Keener) 

Matrix_Week9 <- Skewing_Matrix(Initial_Matrix_Week9)

eigen_Week9 <- eigen(Matrix_Week9)

team_ranking_week9 <- data.frame(Team = c("Bills", "Dolphins", "Patriots", "Jets", "Ravens", "Bengals", "Browns", "Steelers", "Texans", "Colts", "Jaguars", "Titans", "Broncos", "Chiefs", "Raiders", "Chargers", "Cowboys", "Giants", "Eagles", "Commanders", "Bears", "Lions", "Packers", "Vikings", "Falcons", "Panthers", "Saints", "Buccaneers", "Cardinals", "Rams", "49ers", "Seahawks"),
                         Ranking_Value = eigen_Week9$vectors[,1]) |> 
  arrange(-Ranking_Value)

team_ranking_week9
```

## Least Squares Through Week 9

```{r}
Solutions_week9 <- as.matrix(Week_9_Least_Squares[, "home_win_margin"])

zeros_week9 <- matrix(0, ncol = 32, nrow = 138)

T_week9 <- modify_matrix(zeros_week9, Week_9_Least_Squares)

NewT_week9 <- t(T_week9)%*%T_week9

NewSolution_week9 <- t(T_week9)%*%Solutions_week9

LinearIndependence_week9 <- modify_matrix2(NewT_week9, 32)

NewSolution_week9[32, 1] <- 0

NFL_Rankings_Week9 <- solve(LinearIndependence_week9) %*% NewSolution_week9

Ranking_Least_SquaresWeek9 <- data.frame(Team = c("Bills", "Dolphins", "Patriots", "Jets", "Ravens", "Bengals", "Browns", "Steelers", "Texans", "Colts", "Jaguars", "Titans", "Broncos", "Chiefs", "Raiders", "Chargers", "Cowboys", "Giants", "Eagles", "Commanders", "Bears", "Lions", "Packers", "Vikings", "Falcons", "Panthers", "Saints", "Buccaneers", "Cardinals", "Rams", "49ers", "Seahawks"),
                         Ranking_Value = NFL_Rankings_Week9[,1]) |> 
  arrange(-Ranking_Value)

Ranking_Least_SquaresWeek9
```

# Comparing Like Models Week to Week

## Comparing Week 9 to Week 18 Keener's

```{r}
Comparing_Keeners <- Comparison(team_ranking_week9, team_ranking_week18)

Comparing_Keeners

Comparing_Keeners |> 
  count(differs)


Comparing_Keeners_3 <- move_3_comparison(team_ranking_week9, team_ranking_week18)

Comparing_Keeners_3

Comparing_Keeners_3 |> 
  count(moved_3_or_less)
```

## Comparing Week 9 to Week 18 Least Squares

```{r}
Comparing_Least_Squares <- Comparison(Ranking_Least_SquaresWeek9, Ranking_Least_SquaresWeek18)

Comparing_Least_Squares |> 
  count(differs)

Comparing_Least_Squares_3 <- move_3_comparison(Ranking_Least_SquaresWeek9, Ranking_Least_SquaresWeek18)

Comparing_Least_Squares_3

Comparing_Least_Squares_3 |> 
  count(moved_3_or_less)
```

# Comparing Same Weeks Model to Model

## Comparing Keener's Week 9 and Least Squares Week 9
```{r}
Week9_Comparisons <- Comparison(team_ranking_week9, Ranking_Least_SquaresWeek9)

Week9_Comparisons

Week9_Comparisons |> 
  count(differs)

Week9_Comparisons_3 <- move_3_comparison(team_ranking_week9, Ranking_Least_SquaresWeek9)

Week9_Comparisons_3

Week9_Comparisons_3 |> 
  count(moved_3_or_less)
```

## Comparing Keener's Week 18 and Least Sqaures Week 18

```{r}
Week18_Comparisons <- Comparison(team_ranking_week18, Ranking_Least_SquaresWeek18)

Week18_Comparisons

Week18_Comparisons |> 
  count(differs)

Week18_Comparisons_3 <- move_3_comparison(team_ranking_week18, Ranking_Least_SquaresWeek18)

Week18_Comparisons_3

Week18_Comparisons_3 |> 
  count(moved_3_or_less)
```


# Comparing Models With NFL Final Results

## Keener's Method vs NFL End of Season Rankings

### Week 9

```{r}
NFL_Rankings_Keeners_Week9 <- Comparison(NFL_End_Of_Season_Ranking, team_ranking_week9)

NFL_Rankings_Keeners_Week9

NFL_Rankings_Keeners_Week9 |> 
  count(differs)

NFL_Rankings_Keeners_Week9_3 <- move_3_comparison(NFL_End_Of_Season_Ranking, team_ranking_week9)

NFL_Rankings_Keeners_Week9_3

NFL_Rankings_Keeners_Week9_3 |> 
  count(moved_3_or_less)
```

### Week 18 

```{r}
NFL_Rankings_Keeners_Week18 <- Comparison(NFL_End_Of_Season_Ranking, team_ranking_week18)

NFL_Rankings_Keeners_Week18

NFL_Rankings_Keeners_Week18 |> 
  count(differs)

NFL_Rankings_Keeners_Week18_3 <- move_3_comparison(NFL_End_Of_Season_Ranking, team_ranking_week18)

NFL_Rankings_Keeners_Week18_3

NFL_Rankings_Keeners_Week18_3 |> 
  count(moved_3_or_less)
```

## Keener's Method vs Playoff Placing 

### Week 9

```{r}
top_14_ranking_week9 <- team_ranking_week9 |> 
  slice_head(n = 14)

playoff_placing_week9 <- Comparison(NFL_Playoff_Placing, top_14_ranking_week9)

playoff_placing_week9 |> 
  count(differs)

playoff_placing_week9_3 <- move_3_comparison(NFL_Playoff_Placing, top_14_ranking_week9)

playoff_placing_week9_3 |> 
  count(moved_3_or_less)

playoff_placing_week9_3
```

### Week 18

```{r}
top_14_ranking_week18 <- team_ranking_week18 |> 
  slice_head(n = 14)

top_14_ranking_week18

playoff_placing_week18 <- Comparison(NFL_Playoff_Placing, top_14_ranking_week18)

playoff_placing_week18 |> 
  count(differs)

playoff_placing_week18_3 <- move_3_comparison(NFL_Playoff_Placing, top_14_ranking_week18)

playoff_placing_week18_3 |> 
  count(moved_3_or_less)
```


## Least Squares vs NFL End of Season Rankings 

### Week 9

```{r}
NFL_Rankings_Least_Squares_Week9 <- Comparison(NFL_End_Of_Season_Ranking, Ranking_Least_SquaresWeek9)

NFL_Rankings_Least_Squares_Week9

NFL_Rankings_Least_Squares_Week9 |> 
  count(differs)

NFL_Rankings_Least_Squares_Week9_3 <- move_3_comparison(NFL_End_Of_Season_Ranking, Ranking_Least_SquaresWeek9)

NFL_Rankings_Least_Squares_Week9_3

NFL_Rankings_Least_Squares_Week9_3 |> 
  count(moved_3_or_less)
```

### Week 18 

```{r}
NFL_Rankings_Least_Squares_Week18 <- Comparison(NFL_End_Of_Season_Ranking, Ranking_Least_SquaresWeek18)

NFL_Rankings_Least_Squares_Week18

NFL_Rankings_Least_Squares_Week18 |> 
  count(differs)

NFL_Rankings_Least_Squares_Week18_3 <- move_3_comparison(NFL_End_Of_Season_Ranking, Ranking_Least_SquaresWeek18)

NFL_Rankings_Least_Squares_Week18_3

NFL_Rankings_Least_Squares_Week18_3 |> 
  count(moved_3_or_less)
```

## Least Squares vs Playoff Placing 

### Week 9

```{r}
top_14_ranking_least_squares_week9 <- Ranking_Least_SquaresWeek9 |> 
  slice_head(n = 14)

playoff_placing_least_squares_week9 <- Comparison(NFL_Playoff_Placing, top_14_ranking_least_squares_week9)

playoff_placing_least_squares_week9 |> 
  count(differs)

playoff_placing_least_squares_week9_3 <- move_3_comparison(NFL_Playoff_Placing, top_14_ranking_least_squares_week9)

playoff_placing_least_squares_week9_3 |> 
  count(moved_3_or_less)
```

### Week 18

```{r}
top_14_ranking_least_squares_week18 <- Ranking_Least_SquaresWeek18 |> 
  slice_head(n = 14)

top_14_ranking_least_squares_week18

playoff_placing_least_squares_week18 <- Comparison(NFL_Playoff_Placing, top_14_ranking_least_squares_week18)

playoff_placing_least_squares_week18 |> 
  count(differs)

playoff_placing_least_squares_week18_3 <- move_3_comparison(NFL_Playoff_Placing, top_14_ranking_least_squares_week18)

playoff_placing_least_squares_week18_3 |> 
  count(moved_3_or_less)
```