---
title: "320 Final Project"
author: "Belle Schmidt and Rachael Olson"
date: "2025-11-12"
output: html_document
---

```{r}
library(readr)
library(dplyr)
library(pracma)
nfl_team_stats_2002_2024 <- read_csv("C:\\Users\\isabe\\OneDrive\\Big Data Bowl\\TeamRanking\\Schmidt\\nfl_team_stats_2002-2024.csv")
#View(nfl_team_stats_2002_2024)

# nfl_team_stats_2002_2024
```

```{r}
Season24 <- nfl_team_stats_2002_2024 %>% 
  filter(season == 2024)

Season24 <- Season24 %>% 
  mutate(away_win_margin = score_away - score_home,
         home_win_margin = score_home - score_away) 

Thru_Week_6 <- Season24 %>% 
  filter(week < 6) %>% 
  dplyr::select(-season, -week, -date, -time_et)

Thru_Week_6

Away_Stats <- Thru_Week_6 %>% 
  group_by(away) %>% 
  summarise(points_allowed_away = sum(score_home),
            yards_allowed_away = sum(yards_home),
            sacks_away = sum(sacks_num_away),
            turnovers_away = sum(fumbles_away) + sum(interceptions_away),
            num_away_games = n())

Home_Stats <- Thru_Week_6 %>% 
  group_by(home) %>% 
  summarise(points_allowed_home = sum(score_away),
            yards_allowed_home = sum(yards_away),
            sacks_home = sum(sacks_num_home),
            turnovers_home = sum(fumbles_home) + sum(interceptions_home),
            num_home_games = n())


ThruWeek6Totals <- full_join(Away_Stats, Home_Stats, join_by(away == home))

Season24 %>% 
  dplyr::select(-season, -date, -time_et, -first_downs_away, -first_downs_home, -first_downs_from_passing_away, -first_downs_from_passing_home, -first_downs_from_rushing_away, -first_downs_from_rushing_home, -first_downs_from_penalty_away, -first_downs_from_penalty_home, -third_down_comp_away, -third_down_comp_home, -third_down_att_away, -third_down_att_home, -fourth_down_att_away, -fourth_down_att_home, -fourth_down_comp_away, -fourth_down_comp_home, -plays_away, -plays_home, -drives_away, -drives_home, -pass_comp_away, -pass_comp_home)

#View(Season24)
```

# Pre Information

AFC East: 
T1 = Bills
T2 = Dolphins
T3 = Patriots
T4 = Jets

AFC North: 
T5 = Ravens
T6 = Bengals
T7 = Browns
T8 = Steelers

AFC South: 
T9 = Texans
T10 = Colts 
T11 = Jaguars
T12 = Titans

AFC West: 
T13 = Broncos
T14 = Chiefs 
T15 = Raiders
T16 = Chargers

NFC East: 
T17 = Cowboys
T18 = Giants
T19 = Eagles
T20 = Commanders

NFC North: 
T21 = Bears
T22 = Lions
T23 = Packers
T24 = Vikings

NFC South: 
T25 = Falcons
T26 = Panthers
T27 = Saints
T28 = Buccaneers

NFC West: 
T29 = Cardinals
T30 = Rams
T31 = 49ers
T32 = Seahawks

```{r}
Least_Squares_data <- Season24 |> 
  select(week, home, away, home_win_margin) |> 
  mutate(home = ifelse(home == "Bills", "T1", ifelse(home == "Dolphins", "T2", ifelse(home == "Patriots", "T3", ifelse(home == "Jets", "T4", ifelse(home == "Ravens", "T5", ifelse(home == "Bengals", "T6", ifelse(home == "Browns", "T7", ifelse(home == "Steelers", "T8", ifelse(home == "Texans", "T9", ifelse(home == "Texans", "T9", ifelse(home == "Colts", "T10", ifelse(home == "Jaguars", "T11", ifelse(home == "Titans", "T12", ifelse(home == "Broncos", "T13", ifelse(home == "Chiefs", "T14", ifelse(home == "Raiders", "T15", ifelse(home == "Chargers", "T16", ifelse(home == "Cowboys", "T17", ifelse(home == "Giants", "T18", ifelse(home == "Eagles", "T19", ifelse(home == "Commanders", "T20", ifelse(home == "Bears", "T21", ifelse(home == "Lions", "T22", ifelse(home == "Packers", "T23", ifelse(home == "Vikings", "T24", ifelse(home == "Falcons", "T25", ifelse(home == "Panthers", "T26", ifelse(home == "Saints", "T27", ifelse(home == "Buccaneers", "T28", ifelse(home == "Cardinals", "T29", ifelse(home == "Rams", "T30", ifelse(home == "49ers", "T31", "T32"))))))))))))))))))))))))))))))))) |> 
  mutate(away = ifelse(away == "Bills", "T1", ifelse(away == "Dolphins", "T2", ifelse(away == "Patriots", "T3", ifelse(away == "Jets", "T4", ifelse(away == "Ravens", "T5", ifelse(away == "Bengals", "T6", ifelse(away == "Browns", "T7", ifelse(away == "Steelers", "T8", ifelse(away == "Texans", "T9", ifelse(away == "Texans", "T9", ifelse(away == "Colts", "T10", ifelse(away == "Jaguars", "T11", ifelse(away == "Titans", "T12", ifelse(away == "Broncos", "T13", ifelse(away == "Chiefs", "T14", ifelse(away == "Raiders", "T15", ifelse(away == "Chargers", "T16", ifelse(away == "Cowboys", "T17", ifelse(away == "Giants", "T18", ifelse(away == "Eagles", "T19", ifelse(away == "Commanders", "T20", ifelse(away == "Bears", "T21", ifelse(away == "Lions", "T22", ifelse(away == "Packers", "T23", ifelse(away == "Vikings", "T24", ifelse(away == "Falcons", "T25", ifelse(away == "Panthers", "T26", ifelse(away == "Saints", "T27", ifelse(away == "Buccaneers", "T28", ifelse(away == "Cardinals", "T29", ifelse(away == "Rams", "T30", ifelse(away == "49ers", "T31", "T32")))))))))))))))))))))))))))))))))
```

## Massey's Method General: Least Squares

```{r}
# Function That Makes Initial T Matrix
modify_matrix <- function(input_matrix, data) {
  for (i in 1:nrow(data)){
    col <- data$home[i]
    col_index <- as.numeric(sub("T","", data$home[i]))
    input_matrix[i, col_index] <- 1
    col2 <- data$away[i]
    col_index2 <- as.numeric(sub("T","", data$away[i]))
    input_matrix[i, col_index2] <- -1
    
  
  }
  
  return(input_matrix)
}

# Used For Fixing The Linear Dependence Issue
modify_matrix2 <- function(input_matrix, row) {
  for (i in 1:ncol(input_matrix)){
    input_matrix[row, i] <- 1
    
}
  
  return(input_matrix)
    
}
```



# Least Squares Through Week 9

```{r}
Solutions_week9 <- as.matrix(Week_9_Least_Squares[, "home_win_margin"])

zeros_week9 <- matrix(0, ncol = 32, nrow = 138)

T_week9 <- modify_matrix(zeros_week9, Week_9_Least_Squares)

NewT_week9 <- t(T_week9)%*%T_week9

NewSolution_week9 <- t(T_week9)%*%Solutions_week9

LinearIndependence_week9 <- modify_matrix2(NewT_week9, 32)

NewSolution_week9[32, 1] <- 0

NFL_Rankings_Week9 <- solve(LinearIndependence_week9) %*% NewSolution_week9

NFL_Rankings_Week9
```

# Least Squares Ranking Week 18

```{r}
Solutions_week18 <- as.matrix(Least_Squares_data[, "home_win_margin"])

zeros_week18 <- matrix(0, ncol = 32, nrow = 272)

T_week18 <- modify_matrix(zeros_week18, Least_Squares_data)

NewT_week18 <- t(T_week18)%*%T_week18

NewSolution_week18 <- t(T_week18)%*%Solutions_week18

LinearIndependence_week18 <- modify_matrix2(NewT_week18, 32)

NewSolution_week18[32, 1] <- 0

NFL_Rankings_Week18 <- solve(LinearIndependence_week18) %*% NewSolution_week18

NFL_Rankings_Week18
```